---
title: 'Coursera Case Study: Bikes'
author: "Matthew"
date: "2022-11-10"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500)
library(tidyverse)
library(scales)
library(tidymodels)
library(lubridate)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
```

Multiple datasets (13) were downloaded from a link provided by Coursera and then was merged into one csv, which was subsequently read into using the code chunk below. The first chunk uses a for-loop to iterate the *read_csv* (read data files) and *rbind* (combine data rowwise) functions till all data files are read; which is then saved as a csv of its own. Now the csv can be read only once using the following code chunk.

```{r eval=FALSE}
files <- list.files("D:/Downloads/Case Study")
bikes <- read_csv("D:/Downloads/Case Study/202110-divvy-tripdata.csv")

for(i in 2:length(files)) {
  bikes <- bikes %>% rbind(read_csv(paste0("D:/Downloads/Case Study/",
                   files[i])))
}

write_csv(bikes, "bikes.csv")
```

```{r}
bikes <- read_csv("D:/Downloads/Case Study/Full Data/bikes.csv")
```

Create columns that are the *length* of bike ride (in hours) and *day_of_week* (1 = Monday, 7 = Sunday).

```{r}
bikes$length <- difftime(bikes$ended_at, bikes$started_at, "hours")
bikes$day_of_week <- wday(bikes$started_at)
```

```{r}
skimr::skim_without_charts(bikes)
```

