---
title: "NYC House Prices"
author: "Matthew"
date: "4/24/2022"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 600)
library(tidyverse)
library(skimr)
library(patchwork)
library(broom)
library(tidymodels)
library(rnaturalearth)
library(rnaturalearthdata)
library(splines)
theme_set(theme_bw())
```


```{r}
house <- read_csv("C:/Users/Matthew Hondrakis/OneDrive/Documents/data_lat_long2.csv")
house <- house[,-1]
```

# Explore and Clean
```{r}
house[1:20,] %>% 
  View()

skim(house)

house <- 
  house %>% rename_with(tolower) %>% 
  rename_with( ~ gsub(" information", "", .x)) %>% 
  rename_with( ~ gsub(" ", "_", .x))
```


```{r}
house <- house %>% 
  mutate(across(c(bath:sqft, tax:total_cost), ~ parse_number(.x)))

house <- house %>% 
  rename(by_car = commute) %>% 
  mutate(by_car = parse_number(by_car))


house <- house %>% 
  mutate(type = sub(",.*","", home_details))
```


```{r}
house %>% 
  keep(is.numeric) %>% 
  select(-lon, -lat) %>% 
  gather() %>% 
  ggplot(aes(value)) + geom_histogram() + facet_wrap(~key, scales = "free")

house %>% 
  keep(is.numeric) %>% 
  select(-lon, -lat) %>% 
  pivot_longer(-price) %>% 
  ggplot(aes(value, price)) + geom_point() +
  scale_y_log10() + scale_x_log10() +
  facet_wrap(~name)
```


```{r}
house %>% 
  select_if(is.numeric) %>% 
  select(-lon, -lat) %>% 
  drop_na() %>% 
  pivot_longer(-price) %>% 
  group_by(name) %>% 
  summarize(corr = cor(price, value)) %>% 
  arrange(-abs(corr))


tidy_model <- house %>% 
  select_if(is.numeric) %>% 
  select(-lon, -lat) %>% 
  drop_na() %>% 
  pivot_longer(-price) %>% 
  nest(-name) %>% 
  mutate(mod = map(data, ~ lm(log(price) ~ value, .x)),
         tidy = map(mod, broom::tidy)) %>% 
  unnest(tidy)
```


```{r}
tidy_model %>% 
  filter(term == "value") %>% 
  arrange(-abs(estimate))
```

## Checking correlated variables
```{r}
gplot <- function(x){
  (house %>% 
    filter(!is.na({{x}}), !is.na(price)) %>% 
    ggplot(aes({{x}}, price, group = {{x}})) + geom_boxplot() + scale_y_log10()) +
  (house %>% 
     filter(!is.na({{x}}), !is.na(price)) %>% 
     group_by({{x}}) %>% 
     summarize(m = mean(price, na.rm = TRUE)) %>% 
     ggplot(aes({{x}}, m)) + geom_line() + scale_y_log10())
}

gplot(bath)
gplot(bed)
gplot(assessment_year)
```


```{r}
house %>% 
  filter(price != 0, !is.na(price)) %>% 
  mutate(price = price/1e3,
         type = fct_lump(type, 7),
         type = fct_reorder(type, price, median)) %>% 
  ggplot(aes(price, type)) + geom_boxplot() +
  scale_x_log10(labels = scales::comma) + labs(x = "", title = "Price in thousands") +
  theme(plot.margin = margin(10,50,10,0))
```


## Plots of Price by numerics
```{r}
gplot2 <- function(x){
  house %>% 
    ggplot(aes({{x}}, price, color = fct_lump(type, 6))) + 
    geom_point() +
    scale_x_log10() + scale_y_log10()
}
(gplot2(tax) +
gplot2(total_cost) +
gplot2(land_assessment_cost)) /
(gplot2(sqft) +
gplot2(improvement_cost)) + plot_layout(guides = 'collect')
```

# Model
## Preprocess
```{r}
house <- house %>% 
  mutate(type_mod = fct_lump(type, 6))

house_mod <- house %>% 
  select(price, tax, total_cost, sqft, land_assessment_cost,
         improvement_cost, type_mod, bath, bed, address) %>% 
  drop_na() %>% 
  filter_if(is.numeric, all_vars(. > 0))

house_mod %>% 
  keep(is.numeric) %>% 
  cor()
```


```{r}
house_mod %>% 
  keep(is.numeric) %>% 
  pivot_longer(-tax) %>% 
  ggplot(aes(tax, value)) + geom_point() +
  facet_wrap(~name, scales = "free")

house_mod %>% 
  ggplot(aes(bath, price)) + geom_point() + scale_y_log10() +
  geom_smooth()
```


```{r}
updated_model <- 
  lm(log(price) ~ (log(sqft) + log(tax) + bath) * type_mod, 
   house_mod)

anova(updated_model)
summary(updated_model)

(augment(updated_model) %>% 
  ggplot(aes(`log(price)`,.fitted, color = type_mod)) + 
  geom_point(alpha = 0.5) + geom_abline()) /
(augment(updated_model) %>% 
  mutate(residual = .fitted - `log(price)`) %>% 
  ggplot(aes(.fitted, residual, color = type_mod)) + 
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0)) + plot_layout(guides = "collect")

hist(residuals(updated_model))
```

## Tidymodels

### Set up and data split
```{r}
set.seed(123)
house_split <- initial_split(house_mod %>% 
                               mutate(price = log(price),
                                      tax = log(tax),
                                      sqft = log(sqft),
                                      bath = log(bath), 
                             strata = bath))
house_test <- testing(house_split)
house_train <- training(house_split)
```


### Model creating and fit
```{r}
mod <- linear_reg() %>% 
  set_mode("regression") %>% 
  set_engine("lm")
rec <- recipe(price ~ tax + sqft + 
                     bath + type_mod, house_train) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):all_predictors())
wkfl <- workflow() %>% 
  add_model(mod) %>% 
  add_recipe(rec)
wkfl_fit <- fit(wkfl, house_train)
```


### Analysis
```{r}
house_res <- 
  house_test %>% 
  bind_cols(predict(wkfl_fit, house_test))
```


```{r}
(house_res %>%
  mutate(residuals = price - .pred) %>% 
  ggplot(aes(.pred, residuals, color = type_mod)) + 
  geom_point(alpha = 0.5) + geom_hline(yintercept = 0)) /
(house_res %>%  
  ggplot(aes(.pred, price, color = type_mod)) + 
  geom_point(alpha = 0.5) + geom_abline()) + plot_layout(guide = "collect")
```


```{r}
three_metrics <- metric_set(rsq, rmse, mae)

joined_metrics <- three_metrics(augment(updated_model) %>% 
      rename(price = `log(price)`),
    price, .fitted) %>% 
  mutate(model = "1") %>% 
  bind_rows(three_metrics(house_res, price, .pred) %>% 
              mutate(model = "2"))

joined_metrics %>% arrange(desc(.metric))
```


```{r}
joined_metrics %>% 
  ggplot(aes(.estimate, .metric, fill = model)) +
  geom_col(position = "dodge") +
  labs(title = "Base R model (1) vs Tidymodels (2)",
       caption = "Comparison is not equal; the fit methods were done differently. \n
       Did not separate a training/testing dataset for the first model")
```



```{r}
house <- house %>% 
  mutate(zip_code = str_sub(address, - 5, - 1))

x <- house %>% 
  filter(!is.na(price)) %>% 
  group_by(zip_code) %>% 
  summarize(m = median(price)) %>% 
  arrange(-m) %>% 
  head(20) %>% 
  pull(zip_code)
y <- house %>% 
  filter(!is.na(price)) %>% 
  group_by(zip_code) %>% 
  summarize(m = median(price)) %>% 
  arrange(-m) %>% 
  tail(20) %>% 
  pull(zip_code)
  
house %>% 
  select(zip_code, price) %>% group_by(zip_code) %>% 
  filter(!is.na(price), zip_code %in% c(x,y)) %>% 
  group_by(zip_code) %>% 
  ggplot(aes(price, fct_reorder(zip_code, price, median))) + geom_boxplot() +
  scale_x_log10() + labs(y = "", title = "House Prices in NYC by zipcode", x = "Price (log10 scale)")
```

## No Taxes!
```{r}
wkfl2 <- workflow() %>% 
  add_model(mod) %>% 
  add_recipe(recipe(price ~ sqft + 
                     bath + type_mod, house_train) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):all_predictors()))

wkfl_fit2 <- fit(wkfl2, house_train)
```


```{r}
house_res2 <- house_test %>%
  bind_cols(predict(wkfl_fit2, house_test))
  
all_metrics <- joined_metrics %>% 
  bind_rows(three_metrics(house_res2, price, .pred) %>% 
              mutate(model = "3"))
```


```{r}
all_metrics %>% 
  ggplot(aes(.estimate, .metric, fill = model)) +
  geom_col(position = "dodge")

all_metrics %>% 
  arrange(desc(.metric), -.estimate)
```


# Borough?
```{r}
house %>% 
  mutate(borough = case_when(
    str_detect(address, "New York") ~ "New York",
    str_detect(address, "Brooklyn") ~ "Brooklyn",
    str_detect(address, "Queens") ~ "Queens",
    str_detect(address, "Bronx") ~ "Bronx",
    str_detect(address, "Staten Island") ~ "Staten Island"
  )) %>% 
  filter(!is.na(borough), !is.na(price)) %>% 
  ggplot(aes(price, borough)) + geom_boxplot() + scale_x_log10()
```



## Don't miss the forrest for the trees
```{r}
house_mod <- house_mod %>%
  mutate(zip_code = str_sub(address, -5, -1))

house_mod %>% 
  ggplot(aes(tax, price, color = zip_code)) + geom_point() +
  theme(legend.position = "none") + scale_x_log10() +
  scale_y_log10()

house_mod2 <- house_mod %>% 
  mutate(zip_code = fct_lump(zip_code, prop = 0.003))
```


```{r}
set.seed(123)
split2 <- initial_split(house_mod2 %>% 
                               mutate(price = log(price),
                                      tax = log(tax),
                                      sqft = log(sqft),
                                      zip_code = factor(zip_code),
                                      bath = factor(bath)),
                        strata = zip_code)


train2 <- training(split2)
test2 <- testing(split2)
```


```{r}
modrf <- rand_forest() %>% 
  set_mode("regression") %>% 
  set_engine("ranger")


recrf <- recipe(price ~ sqft + type_mod + zip_code, train2) %>% 
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~ starts_with("type"):all_predictors())

wkflrf <- workflow() %>% 
  add_model(modrf) %>% 
  add_recipe(recrf)
```


```{r}
wkflrf_fit <- fit(wkfl2, train2)

house_res2 <- 
  test2 %>% 
  bind_cols(predict(wkflrf_fit, test2))

all_metrics_rf <- all_metrics %>% 
  bind_rows(three_metrics(house_res2, price, .pred) %>% 
              mutate(model = "4"))

all_metrics_rf %>% 
  arrange(desc(.metric), -.estimate)
```


```{r}
all_metrics_rf %>% 
  ggplot(aes(.estimate, .metric, fill = fct_reorder(model, .estimate, max, .desc = TRUE))) +
  geom_col(position = "dodge") + 
  labs(title = "All 4 Models", 
       subtitle = "Base R lm (Red), TM lm (Green), TM no tax lm (Blue), Random Forrest (Purple)") +
  theme(legend.position = "none") + scale_x_continuous(breaks = seq(0.3,0.7,0.05))
```



# Just let Gam figure it out
```{r}
library(mgcv)
library(gratia)
set.seed(10)
gam_data <- house %>% 
  select(price, tax, sqft, bath, lon, lat, type_mod) %>% 
  mutate(price = log(price)) %>% 
  drop_na()

gam_split <- initial_split(gam_data, strata = type_mod)

gam_train <- training(gam_split)
gam_test <- testing(gam_split)

gam_mod <- gam(price ~ 
                 s(sqft, by = type_mod) + 
                 s(bath, by = type_mod) + 
                 s(lon, lat), 
               gaussian, gam_train, method = 'REML')

summary(gam_mod)

appraise(gam_mod)

draw(gam_mod, select = smooths(gam_mod)[1:6])

draw(gam_mod, select = smooths(gam_mod)[7:12])

draw(gam_mod, select = smooths(gam_mod)[13])

augment(gam_mod) %>% 
  three_metrics(price, .fitted)

(gam_test %>% 
  mutate(predict = predict(gam_mod, gam_test)) %>% 
  ggplot(aes(price, predict, color = type_mod)) + geom_point(alpha = 0.3) +
  geom_abline()) /
(gam_test %>% 
  mutate(predict = predict(gam_mod, gam_test),
         resid = price - predict) %>% 
  ggplot(aes(predict, resid, color = type_mod)) + geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0)) + plot_layout(guides = "collect")

k.check(gam_mod)
```

I would correct the 'k' value for the bath and lon/lat variables, but it took too long for my computer to run. So, we will leave it like this!

And if you are wondering, its metrics did improve.



```{r}
all_metrics_rf %>% 
  bind_rows(augment(gam_mod) %>% 
    three_metrics(price, .fitted) %>% mutate(model = "5")) %>% 
  arrange(desc(.metric), -.estimate)

all_metrics_rf %>% 
  arrange(desc(.metric), -.estimate) %>% 
  bind_rows(augment(gam_mod) %>% 
  three_metrics(price, .fitted) %>% 
    mutate(model = "5")) %>% 
  ggplot(aes(.estimate, .metric, fill = fct_reorder(model, .estimate, max, .desc = TRUE))) +
  geom_col(position = "dodge") + 
  labs(title = "All 5 Models", 
       subtitle = "Base R lm (Brownish-Yellow), TM lm (Green), TM no tax lm (Blue), \n
       Random Forrest (Purple), Gam (Red)") +
  theme(legend.position = "none")
```



# Scrap work
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

range(house$lon, na.rm = TRUE)
range(house$lat, na.rm = TRUE)

ggplot(world) + geom_sf() +
  coord_sf(
    xlim = c(-74.35, -73.6), 
    ylim = c(min(house$lat, na.rm = TRUE), 40.95), expand = FALSE) + 
  geom_point(data = house %>% filter(!is.na(price)) %>% mutate(price = log(price)), 
             aes(lon, lat, color = price, alpha = 0.3)) +
  scale_color_viridis_c()
```

Because of the lack of precision in longitude/latitude variables, the points on the map are slightly off.


```{r}
house %>% 
  filter(!is.na(price), between(lon, -75, -73), between(lat, 40, 41)) %>% 
  select(price, lon, lat) %>% 
  pivot_longer(-price) %>% 
  ggplot(aes(value, log(price))) + geom_point() + geom_smooth(formula = y ~ ns(x, 5)) +
  facet_wrap(~name, scales = "free")
```


