---
title: "Stock Prices of Popular Tech Companies"
author: "Matthew"
date: "2023-02-02"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500, fig.showtext = TRUE)
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
library(sysfonts)
library(showtext)
library(corrr)
library(rstatix)
font_add_google("Roboto Slab", "Roboto")
theme_set(theme_minimal())
theme_update(plot.title = element_text(hjust = 0.5),
             text = element_text(family = "Roboto"))
```

*Data from Evan Gower on [Kaggle](https://www.kaggle.com/datasets/evangower/big-tech-stock-prices?resource=download&select=TSLA.csv)*

# Read Data

```{r message=FALSE}
files <- list.files(path = "./Data")

stocks <- read_csv(paste0("./Data/", files), 
                   id = "name") %>% 
  mutate(name = gsub("./Data/", "", name),
         name = gsub("\\.csv","",name)) %>% 
  rename_with(tolower)
```

I'm reading in a list of *csv* files from a directory and creating a dataframe from them. I'm then cleaning up the "name" column in the dataframe and storing the cleaned "name" column in the dataframe as the final output.

```{r message=FALSE}
big_tech_companies <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-07/big_tech_companies.csv')

stocks <- stocks %>% 
  inner_join(big_tech_companies, by = c("name" = "stock_symbol"))
```

Datasets are joined in order to retrieve the full names of the companies.

# Exploratory Data Analysis

```{r}
end_labels <- 
  (stocks %>% 
  group_by(company) %>% 
  filter(date == max(date)) %>% 
  arrange(-open) %>% 
  select(open, company))[c(1:3,12:14),]

stocks %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  scale_y_continuous(sec.axis = sec_axis(~., breaks = end_labels$open, 
                        labels = end_labels$company)) +
  scale_x_date(expand = c(0,0)) +
  labs(x = "", y = "Open", color = "",
       title = "Opening Stock Prices of Major Tech Companies",
       subtitle = "Prices range from Jan 2010 to Jan 2023. The names shown are\nthe top and bottom 3 tech stocks at the most recent date.") +
  theme(legend.position = "none")
```

## Top 3

```{r}
(stocks %>% 
  filter(company %in% end_labels$company[1:3]) %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  facet_wrap(~company) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Top 3", x = "")) /
(stocks %>% 
  filter(company %in% end_labels$company[-(1:3)]) %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  facet_wrap(~company) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Bottom 3", x = ""))
```

## Peak Month

```{r}
stocks %>% 
  group_by(company) %>% 
  slice_max(open) %>% 
  filter(year(date) > 2020) %>% 
  mutate(month = month(date, label = TRUE),
         month = fct_relevel(month, "Jan", after = 12),
         month = fct_relevel(month, "Feb", after = 12)) %>% 
  ggplot(aes(month)) +
  geom_bar(color = "black", fill = "steelblue2") +
  geom_vline(xintercept = 5.5, linetype = "dashed") +
  geom_text(label = "2022", x = 5.8, y = 5, 
            color = "steelblue2", fontface = "bold") +
  geom_text(label = "2021", x = 5.2, y = 5,
            color = "steelblue2", fontface = "bold") +
  labs(title = "Peak Month for Tech Stocks", x = "", y = "Number of Stocks",
       subtitle = "Out of 13 tech stocks, 6 of them reached their peak in November 2021",
       caption = "IBM's peak was in March, 2013 and thus is not included.")
```

## Volume

```{r}
stocks %>% 
  mutate(diff = close - open) %>% 
  ggplot(aes(volume, diff)) +
  geom_point(alpha = 0.2) +
  scale_x_log10()

cor((stocks %>% mutate(diff = close - open))$diff, stocks$volume)
```

# Time Series

```{r}
library(prophet)
library(astsa)
library(forecast)
```

## Apple

```{r}
aapl <- stocks %>% 
  filter(name == "AAPL") %>% 
  select(ds = date, y = open)
```

Often, the differences allows us to understand how volatile a trend is. As can be seen below, the **Apple** stock becomes increasingly more volatile over time.

```{r}
(aapl %>% 
  mutate(diff = c(NA, diff(y))) %>% 
  ggplot(aes(ds, diff)) + 
  geom_point(color = "steelblue4", alpha = 0.7) +
  labs(y = "Difference", x = "Date",
       title = "One Day Returns")
) /
(aapl %>% 
   mutate(diff = c(NA, diff(y))) %>% 
   ggplot(aes(diff)) +
   geom_histogram(bins = 50, fill = "steelblue4", color = "black")
)
```

### Prophet

```{r}
m_aapl <- prophet(aapl)

forecast <- predict(m_aapl, 
                    make_future_dataframe(m_aapl, periods = 140))

plot(m_aapl, forecast)

prophet_plot_components(m_aapl, forecast)
```

### ARIMA

First, we begin with fitting an "auto.arima", which automatically fits an ARIMA model with by choosing optimal ARIMA terms.

```{r}
ts_aapl <- ts(aapl$y, start = c(2010, 4), frequency = 365)
aapl_fit <- window(ts_aapl, end = 2018)
```

```{r}
auto_arima_fit <- auto.arima(aapl_fit)
plot(forecast(auto_arima_fit, h = 365), ylim = c(0,200))
lines(ts_aapl)
```

Clearly, we can see that an ARIMA(5,2,0) expects it to keep increasing. This is because of the 5 AR terms, strongly predicting it will keep the same trend it had. Now, we will try to fit an ARIMA by looking at a stationary version of the time series. It appears as though the best fit in an ARIMA(2,1,3), although this is of course not the true model for the stock. ARIMA models are not great at modeling stock prices, as their prices are often a product of unknown variables.

*I have removed the diagnostic plots from the output of this markdown file.*

```{r}
plot(diff(log(aapl$y)),type = "l")
acf2(diff(log((aapl$y))))
```

```{r}
sarima.for(aapl_fit, 
           n.ahead = 365,
           2,1,3, 
           plot.all = TRUE)$plot
lines(ts_aapl)
```

# Correlated Stocks

```{r}
(stock_corr <- stocks %>% 
  widyr::pairwise_cor(company, date, open) %>% 
  filter(item1 > item2) %>% 
  mutate(corrstr = ifelse(abs(correlation > 0.5), "Strong", "Weak"),
         type = ifelse(correlation > 0, "Positive", "Negative")) %>% 
  arrange(-abs(correlation)))
```

```{r}
stock_corr %>% 
  ggplot(aes(correlation)) +
  geom_histogram(aes(fill = type), 
                 alpha = 0.7, binwidth = 0.05) +
  xlim(c(-1,1)) +
  labs(title = "Distribution of Correlation Values",
       subtitle = "The majority of companies have a strong positive correlation",
       fill = "Positive Correlation") +
  theme(legend.position = c(0.2,0.8),
        legend.background = element_rect(fill = "white", color = "white"))
```

## IBM, The Outlier

```{r}
(stocks %>% 
  widyr::pairwise_cor(name, date, open) %>% 
  rename(var1 = item1, var2 = item2) %>% 
  cor_spread(value = "correlation") %>% 
  rename(term = rowname))[c(14,1:13),] %>% 
  network_plot()
```

## Positively Correlated

```{r}
stocks %>% 
  filter(company %in% 
           c(stock_corr[1:5,]$item1, stock_corr[1:5,]$item2)) %>% 
  ggplot(aes(date, open, color = company)) + 
  geom_line() +
  labs(x = "", y = "Open Price", color = "",
       title = "The 6 Most Correlated Stocks Have Nearly Identical Trends") +
  theme(legend.position = c(0.2,0.75),
        legend.background = element_rect(fill = "white",
                                         color = "white"))

stocks %>% 
  filter(company %in% c(stock_corr[1,1:2])) %>% 
  select(date, company, open) %>% 
  pivot_wider(names_from = company, values_from = open) %>% 
  ggplot(aes(`Adobe Inc.`, `Amazon.com, Inc.`)) +
  geom_point(alpha = 0.7, color = "steelblue2") +
  geom_smooth(method = "lm", se = FALSE, color = "black",
              linetype = "dashed") +
  labs(title = "Amazon and Adobe Trend")
```

## Negatively Correlated

```{r}
stock_corr %>% 
  filter(str_detect(item1, "Netflix") & str_detect(item2, "Machine"))

stocks %>% 
  filter(str_detect(company, "Netflix|Machine")) %>% 
  ggplot(aes(date, open, color = name)) + 
  geom_line() +
  labs(x = "", y = "Open Price", color = "",
       title = "IBM and Netflix Have Very Different Trends") +
  theme(legend.position = c(0.45,0.8))

stocks %>% 
  filter(str_detect(company, "Netflix|Machine")) %>% 
  select(date, name, open) %>% 
  pivot_wider(names_from = name, values_from = open) %>% 
  ggplot(aes(`IBM`, `NFLX`)) +
  geom_point(alpha = 0.7, color = "steelblue2") +
  geom_smooth(method = "lm", se = FALSE, color = "black",
              linetype = "dashed") +
  labs(title = "IBM and Netflix")
```
