---
title: "Stock Prices of Popular Tech Companies"
author: "Matthew"
date: "2023-02-02"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500, fig.showtext = TRUE)
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
library(sysfonts)
library(showtext)
font_add_google("Roboto Slab", "Roboto")
theme_set(theme_minimal())
theme_update(plot.title = element_text(hjust = 0.5),
             text = element_text(family = "Roboto"))
```

*Data from Evan Gower on [Kaggle](https://www.kaggle.com/datasets/evangower/big-tech-stock-prices?resource=download&select=TSLA.csv)*

# Read Data

```{r message=FALSE}
files <- list.files(path = "./Data")

stocks <- read_csv(paste0("./Data/", files), 
                   id = "name") %>% 
  mutate(name = gsub("./Data/", "", name),
         name = gsub("\\.csv","",name)) %>% 
  rename_with(tolower)
```

I'm reading in a list of *csv* files from a directory and creating a dataframe from them. I'm then cleaning up the "name" column in the dataframe and storing the cleaned "name" column in the dataframe as the final output.

```{r message=FALSE}
big_tech_companies <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-07/big_tech_companies.csv')

stocks <- stocks %>% 
  inner_join(big_tech_companies, by = c("name" = "stock_symbol"))
```

Datasets are joined in order to retrieve the full names of the companies.

# Exploratory Data Analysis

```{r}
end_labels <- 
  (stocks %>% 
  group_by(company) %>% 
  filter(date == max(date)) %>% 
  arrange(-open) %>% 
  select(open, company))[c(1:3,12:14),]

stocks %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  scale_y_continuous(sec.axis = sec_axis(~., breaks = end_labels$open, 
                        labels = end_labels$company)) +
  scale_x_date(expand = c(0,0)) +
  labs(x = "", y = "Open", color = "",
       title = "Opening Stock Prices of Major Tech Companies",
       subtitle = "Prices range from Jan 2010 to Jan 2023. The names shown are\nthe top and bottom 3 tech stocks at the most recent date.") +
  theme(legend.position = "none")
```

## Top 3

```{r}
(stocks %>% 
  filter(company %in% end_labels$company[1:3]) %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  facet_wrap(~company) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Top 3", x = "")) /
(stocks %>% 
  filter(company %in% end_labels$company[-(1:3)]) %>% 
  ggplot(aes(date, open)) +
  geom_line(aes(color = company)) +
  facet_wrap(~company) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Bottom 3", x = ""))
```

## Peak Month

```{r}
stocks %>% 
  group_by(company) %>% 
  slice_max(open) %>% 
  filter(year(date) > 2020) %>% 
  mutate(month = month(date, label = TRUE),
         month = fct_relevel(month, "Jan", after = 12),
         month = fct_relevel(month, "Feb", after = 12)) %>% 
  ggplot(aes(month)) +
  geom_bar(color = "black", fill = "steelblue2") +
  geom_vline(xintercept = 5.5, linetype = "dashed") +
  geom_text(label = "2022", x = 5.8, y = 5, 
            color = "steelblue2", fontface = "bold") +
  geom_text(label = "2021", x = 5.2, y = 5,
            color = "steelblue2", fontface = "bold") +
  labs(title = "Peak Month for Tech Stocks", x = "", y = "Number of Stocks",
       subtitle = "Out of 13 tech stocks, 6 of them reached their peak in November 2021",
       caption = "IBM's peak was in March, 2013 and thus is not included.")
```

## Volume

```{r}
stocks %>% 
  mutate(diff = close - open) %>% 
  ggplot(aes(volume, diff)) +
  geom_point(alpha = 0.2) +
  scale_x_log10()

cor((stocks %>% mutate(diff = close - open))$diff, stocks$volume)
```

# Time Series

```{r}
library(prophet)
library(astsa)
library(forecast)
```

```{r}
aapl <- stocks %>% 
  filter(name == "AAPL") %>% 
  select(ds = date, y = open)
```

## Apple

Often, the differences allows us to understand how volatile a trend is. As can be seen below, the **Apple** stock becomes increasingly more volatile over time.

```{r}
(aapl %>% 
  mutate(diff = c(NA, diff(y))) %>% 
  ggplot(aes(ds, diff)) + 
  geom_point(color = "steelblue4", alpha = 0.7) +
  labs(y = "Difference", x = "Date",
       title = "One Day Returns")
) /
(aapl %>% 
   mutate(diff = c(NA, diff(y))) %>% 
   ggplot(aes(diff)) +
   geom_histogram(bins = 50, fill = "steelblue4", color = "black")
)
```

### Prophet

```{r}
m_aapl <- prophet(aapl)
future_dates <- make_future_dataframe(m_aapl, periods = 90)

forecast <- predict(m_aapl, future_dates)

plot(m_aapl, forecast)

prophet_plot_components(m_aapl, forecast)
```

### ARIMA

First, we begin with fitting an "auto.arima", which automatically fits an ARIMA model with by choosing optimal ARIMA terms.

```{r}
arima_fit <- auto.arima(aapl$y)
plot(forecast(arima_fit, h = 100))
```

```{r}
plot(diff(log(aapl$y)),type = "l")
acf2(diff(log((aapl$y))))

sarima(log(aapl$y), 2,1,3)
```

```{r}
ts_aapl <- ts(aapl$y, start = c(2010, 4), frequency = 365)
aapl_fit <- window(ts_aapl, end = 2018)
sarima.for(aapl_fit, 
           n.ahead = 365,
           2,1,3, 
           plot.all = TRUE)
lines(ts_aapl)
```
