---
title: "Masters Project: Job Placement"
author: "Matthew"
date: "2022-12-07"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, dpi = 500)
library(tidyverse)
library(scales)
library(patchwork)
library(tidymodels)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
theme_update(panel.grid.minor = element_blank())
```

```{r}
fall <- read_csv("C:/Users/Matthew Hondrakis/OneDrive/Documents/DataAnalysis/Masters Project Fall Placement/fall2022Placement.csv")
```

# Intro

This is a structured analysis guided by the assignment provided by my friends professor. The task is as follows.

-   **Cleanse**: Look for null data in the table, and use the appropriate strategy to handle null data for each column. Explain why this was the strategy that you used.

-   **Outliers**: Search the data for outliers and remove them. Use the appropriate outlier method(s). Show all work.

-   **Who are these graduates?**: Create 4-5 visualizations providing useful and explanatory information about these graduates. The visualizations should show trends, correlations, and other useful patterns that the data provides.\
    Remember, we are dealing with senior management. The charts must be readable, meaningful and at a summary level.\
    Use at least one group/by or bin in your analysis.

-   **Who are most likely to get placed?**: Create visualizations to show management qualities specific to those who are likely to get placed at jobs. Show three meaningful visualizations.\
    Use at least one group/by or bin in your analysis.

-   **Prepare the data for a regression analysis**: The XYZ management company team has heard that it might be possible to use their data to make predictions. They don't know much about data analytics. We are hoping that an example of what is possible will help them understand better.\
    Prepare the data file to run with a regression analysis. Use the techniques and methods discussed in class. (*I'm not a student, so I'm unfamiliar with what methods were discussed in class*)

-   **Run the regression analysis**: Show the regression scoring. Split the data into training and test sets. Show a confusion matrix for both.

## Data Dictionary

```{r data_dictionary}
knitr::kable(readxl::read_excel("C:/Users/Matthew Hondrakis/OneDrive/Documents/DataAnalysis/Masters Project Fall Placement/Data dictionary.xlsx"))
```

# Cleanse

```{r skim_NA}
skimr::skim(fall) %>% select(skim_type, skim_variable, n_missing)
```

As we can see, we have *N/A* values for:

-   *hsc_s* (Specialization in Higher Secondary Education): **1**

-   *specialisation* (Area of speciality): **1**

-   *hsc_p* (Higher Secondary Education percentile): **2**

-   *salary* (Salary of job offered): **67**

```{r count_NA}
fall %>% 
  count(hsc_s, sort = TRUE)
fall %>% 
  count(specialisation, sort = TRUE)
```

I am going to make a few assumptions. I will take "Mkt" to mean "Marketing", "Fin" to mean "Finance", "HR" to mean "Human Resources". We see only one variable having the value of "Mt", which we may assume is a typo or just an outlier. For simplicity's sake, I will consider this a typo and convert it to "Mkt".

```{r}
fall <- fall %>% 
  mutate(specialisation = 
           ifelse(grepl("Mt", specialisation), "Mkt&HR", specialisation))
```

```{r mutual_NAs}
fall %>% 
  filter(is.na(hsc_s)) %>% 
  select(salary)
fall %>% 
  filter(is.na(specialisation)) %>% 
  select(salary)
```

The row that has a missing value for *hsc_s* (Specialization in Higher Secondary Education) also contains a missing value for *salary*; this is not the case for when specialisation is NA.

```{r}
fall %>% 
  group_by(specialisation) %>% 
  count(hsc_s) %>% 
  mutate(prop = n/sum(n)) %>% 
  arrange(specialisation, -n)
fall %>% 
  group_by(hsc_b) %>% 
  count(hsc_s) %>% 
  mutate(prop = n/sum(n)) %>% 
  arrange(hsc_b, -n)
```

Most of "Mkt&HR" are evenly distributed between **Science** and **Commerce**. Thus its not possible at the moment to decide which one of the 2 should be used to fill in the missing value. With regards to *hsc_b* (Board of Education), the majority of **Central** are in **Commerce**. This may provide a clue later on, but it is still too early to make an assumption.

```{r NA_salary}
fall %>% 
  group_by(missing = is.na(salary)) %>% 
  count(status) %>% 
  arrange(status)
```

As we can see, all missing values for *salary* are from individuals that were not offered a job. Since the Data Dictionary defines *Salary* as "Salary of job offered", individuals that were not offered a job, do not have a salary from an offer. Therefore, *salary* will be 0 for individuals that were not offered a job. We will then drop any rows that had missing information, which in total is only **4**.

```{r}
clean_fall <- fall %>% 
  mutate(salary = ifelse(is.na(salary), 0, salary)) %>% 
  drop_na()
```

# Outliers

```{r}
(fall %>% 
  ggplot(aes(salary)) +
  geom_histogram() +
  scale_x_log10(labels = comma_format())) /
(fall %>% 
  ggplot(aes(salary)) +
  geom_boxplot() +
  scale_x_log10(labels = comma_format()) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        panel.grid.major.y = element_blank())) +
  plot_annotation(title = "Salary Distribution")
```

Removing outliers below by filtering out the top 8 salaries and the lowest salary. We are now left with **202** rows out of **215**.

```{r}
clean_fall %>% 
  filter(salary != 0) %>% 
  arrange(salary) %>% 
  slice(1) %>% 
  pull(salary)

updated_fall <- clean_fall %>% 
  arrange(-salary) %>% 
  slice(-c(1:8)) %>% 
  filter(salary != 45000) 
```

# Who are these graduates?

A quick check on the **correlations** between **numeric** variables.

```{r}
updated_fall %>% 
  keep(is.numeric) %>%
  cor() %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "item1") %>% 
  gather(key = item2, value = corr, -item1) %>% 
  filter(item1 > item2) %>% 
  arrange(-abs(corr))
```

Below I will create a function to quickly plot **counts** of factor variables. It is often advised to make a function if one foresees using the same plot/code often. Since this dataset contains many **factor** variables, I assume there will be a lot of factor count plotting.

```{r count_plot_function}
count_plot <- function(x){
  updated_fall %>% 
    filter(!is.na({{x}})) %>% 
    count({{x}}) %>% 
    ggplot(aes(n, fct_reorder({{x}}, n))) + 
    geom_col(color = "black", fill = "lightblue") +
    geom_text(aes(label = n), hjust = 1.3, size = 3) +
    theme(panel.grid.major.y = element_blank())
}
```

## Count Plots

```{r character_plot, echo=FALSE}
# Plot 1
(count_plot(gender) +
  labs(x = "", y = "", title = "Number of Graduates by Gender")) /
(count_plot(degree_t) +
  labs(y = "", x = "", title = "Degree Type"))

# Plot 2
(count_plot(hsc_s) +
   labs(y = "", x = "", title = "Higher Education Specialisation")) /
(count_plot(workex) +
   labs(y = "", x = "", title = "Work Experience"))

# Plot 3
(count_plot(ssc_b) +
    labs(y = "", x = "", title = "Board of Education")) /
(count_plot(hsc_b) +
   labs(y = "", x = "", title = "Higher Education Board of Education"))

# Plot 4
(count_plot(specialisation) +
    labs(y = "", x = "", title = "Specialization")) /
(count_plot(status) +
   labs(y = "", x = "", title = "Status"))
```

## Gender

### 1. Salary Distribution by Gender

```{r}
updated_fall %>% 
  filter(status == "Placed") %>% 
  ggplot(aes(salary, gender, color = gender)) + 
  geom_boxplot() +
  scale_x_log10(label = comma_format()) +
  labs(y = "", x = "Salary", title = "Salary Distribution by Gender") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank())
```

### 2. Percent Specialization in Higher Secondary Education by Gender

Given the unequal distribution of the *gender* variable, I am plotting the **percent** distribution of each specialization for each gender. As we can see, the percentage of males or females choosing a certain specialization is approximately equal.

```{r}
group_count <- function(x,y){
  updated_fall %>%
    group_by({{x}}) %>% 
    count({{y}}) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(prop, {{x}}, 
               fill = fct_reorder2({{y}}, {{x}}, prop, .desc = FALSE))) +
    geom_col(color = "black", position = position_dodge()) +
    labs(y = "", x = "", fill = "") +
    geom_text(aes(label = paste0(round(prop*100,1), "%")), 
            position = position_dodge2(0.9), hjust = 1, size = 3.5)
}
```

```{r}
group_count(gender,hsc_s) +
  ggtitle("Percent Specialization in Higher Secondary Education by Gender") +
  scale_x_continuous(label = percent_format(), breaks = seq(0,0.5,0.1)) +
  theme(panel.grid.major.y = element_blank())
```

A *chisq.test* is provided below to reinforce our confidence in this assumption. As can be seen, the *p.value* is **0.3856** which means there is not enough evidence to suggest that the two genders have uneven specialization distributions.

```{r}
chisq.test(updated_fall$gender, updated_fall$hsc_s)
```

### 3. Work Experience by Gender

```{r}
group_count(gender, workex) +
  ggtitle("Work Experience by Gender") +
  scale_x_continuous(label = percent_format())
```

## Work Experience

### 4. Work Experience by Undergrad Degree

```{r}
group_count(degree_t, workex) +
  scale_fill_discrete(direction = -1) +
  ggtitle("Work Experience by Undergrad Degree") +
  scale_x_continuous(label = percent_format())
```

The group with the highest percentage of individuals that have *work experience* is **Sci&Tech**; yet every group still have less than **50%** work experience. It may be interesting to see later on how *work experience* relates to *status* (Job Placement), and whether this gives Sci&Tech an advantage.

### 5. Work Experience by Specialization

```{r}
group_count(specialisation, workex) +
  scale_fill_discrete(direction = -1) +
  scale_x_continuous(label = percent_format()) +
  ggtitle("Work Experience by Specialization")
```

There appears to be a difference between the two groups (Finance **vs** HR) with regards to *work experience*, although it does not appear to be statistically significant. For now, while we can't reject the hypothesis that the two groups are equal, we will keep in mind that they do in fact have slightly different proportions.

```{r}
chisq.test(updated_fall$workex, updated_fall$specialisation)
```

## Highschool and Undergrad

### 6. Highschool Specialization and Undergrad Degree

```{r}
updated_fall %>%
  group_by(hsc_s) %>% 
  count(degree_t) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(prop, hsc_s, 
             fill = fct_reorder2(degree_t, hsc_s, prop, .desc = FALSE))) +
  geom_col(color = "black", position = position_dodge()) +
  labs(y = "", x = "", fill = "", 
       title = "Undergrad Degree by Highschool Specialization") +
  geom_text(aes(label = paste0(round(prop*100,1), "%"),
                hjust = ifelse(prop > 0.1, 1.1, -0.2)), 
            position = position_dodge2(0.9), size = 3.5) +
  scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +
  theme(legend.position = "bottom") +
  scale_x_continuous(label = percent_format())
```

We can see that there is a bias with regards to what students study for their Undergrad. Students that specialized in *Science* or *Commerce*, stayed within those fields. We can see that students that studied *Commerce* were overwhelming more likely (**96.3%**) to go onto studying *Comm&Mgmt.* Individuals that studied *Science* were more likely (**59.5%**) to go onto studying *Sci&Tech*. On the other hand, most of the *Art* students (**63%**) later went on to studying *Comm&Mgmt*.

# Who are most likely to get placed?

## Work Experience

```{r}
group_count(workex, status) +
  scale_x_continuous(label = percent_format()) +
  ggtitle("Placement by Work Experience")
```

As expected, individuals that have *work experience* are more likely to get a job offer. Almost **85%** of individuals with work experience, received a job offer.

```{r}
updated_fall %>% 
  ggplot(aes(etest_p, status, color = status)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, height = 0.1) +
  labs(y = "", x = "Employability Test Percentile", 
       title = "Employability Test Percentile by Job Placement")
```

We can see that the distribution of *employability test* scores in the **0** to **25** percentile of *Placed* and *Not Placed* are equivalent. Changes start to show after the **25th** percentile and above. This suggests that fewer students among the higher percentiles were not being placed (offered jobs).

```{r}
updated_fall %>%
  group_by(status) %>% 
  summarize(Quantile = paste0(c(0,25,59,75,100),"%"),
            prob = quantile(etest_p, probs = c(0,.25,.50,.75,1))) %>% 
  pivot_wider(names_from = status, values_from = prob) %>% 
  knitr::kable()
```

## Education

```{r}
(group_count(degree_t, status) +
  ggtitle("Under Graduation Degree by Job Offer") +
  scale_x_continuous(label = percent_format())) /
(group_count(hsc_s, status) +
   ggtitle("Higher Secondary Education by Job Offer") +
   scale_x_continuous(label = percent_format())) +
plot_layout(guides = "collect")
```

With regards to Undergrad degree, both *Sci&Tech* and *Comm&Mgmt* had approximately equal proportions of students getting job offers while *Others* had a majority not getting job offers. As we saw from **Plot 6** *(Highschool Specialization by Undergrad Degree),* **Science** and **Sci&Tech** are correlated, as are **Commerce** and **Comm&Mgmt**.

```{r}
updated_fall %>% 
  group_by(hsc_s, degree_t) %>% 
  count(status, sort = TRUE) %>% 
  ungroup(hsc_s) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(status == "Placed") %>% 
  ggplot(aes(prop, fct_reorder(degree_t, prop, .fun = max, .desc = FALSE), 
             fill = hsc_s)) +
  geom_col(position = position_dodge2(), color = "black") +
  geom_text(aes(label = paste0(round(prop*100,1), "%"),
                hjust = ifelse(prop > 0.1, 1.1, -0.2)),
            position = position_dodge2(0.9)) +
  scale_fill_manual(values = c("pink", "lightblue", "lightgreen")) +
  labs(y = "Undergrad", x = "", fill = "Highschool",
       title = "Percent Job offers by Education") +
  scale_x_continuous(label = percent_format())
```

## Test Scores

```{r}
updated_fall %>% 
  ggplot(aes(etest_p, status, color = status)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, height = 0.1) +
  labs(y = "", x = "Employability Test Percentile", 
       title = "Employability Test Percentile by Job Placement")
```

We can see that the distribution of *employability test* scores in the **0** to **25** percentile of *Placed* and *Not Placed* are equivalent. Changes start to show after the **25th** percentile and above. This suggests that fewer students among the higher percentiles were not being placed (offered jobs).

```{r}
updated_fall %>%
  group_by(status) %>% 
  summarize(Quantile = paste0(c(0,25,59,75,100),"%"),
            prob = quantile(etest_p, probs = c(0,.25,.50,.75,1))) %>% 
  pivot_wider(names_from = status, values_from = prob) %>% 
  knitr::kable()
```
